\section*{Acknowledgments}

%Professor Geoffrey Bicknell
%
%Dr Alexander Wagner
%
%Dr Ralph Sutherland
%
%Professor Brian McNamara
%
%Members of the Mondays regular group meeting
%
%Parents
%
%Brother
%
%Sisters
%
%Shaila 
%
%Emdad Sanjay
%
%Sysman
%
%NCI help
%
%Fellow students of Mount Stromlo


Foremost, I would like to thank my supervisor, Professor Geoffrey Bicknell, for introducing me into the field of theoretical and computational astrophysics. His support and guidence, teaching in a very understandable way, patience, kind heartedness, and most of all his faith in me made this MPhil thesis possible. He is a constant source of inspiration to me. I am honored to have such a great personality as my supervisor.I would like to thank my advisor Dr Alexander Wagner for his immense support in my learning of astronomy and computational physics, especially, the computational codes FLASH and PLUTO. His encouragement always helped me to overcome any frustration related to my study. 
I would like to thank my other advisors, Dr Ralph Sutherland and Professor Brian McNamara for their many suggestions related to my thesis. I am grateful to Professor Gregory Taylor for providing me the 6cm VLA data of the Hydra A galaxy. I would like to thank my sisters and brother for their endless love and support. Thank you Shaila for helping me in English grammar. My life is enriched with all your support, care and love. 

Thanks to my dearest friend Emdad who makes my life in Canberra very joyful. Thanks to my other friends in Australia-- Harun, Sanjay and Aysha. I would like to thank RSAA computer staff for helping me in issues related to my mac laptop, cluster computer mosura or any software.

Thanks to fellow students of RSAA and all academics here. I feel privileged to live and work at Mount Stromlo.I would like to thank NCI system administration staff for their support in the usage of the supercomputers VAYU and RAIJIN. I gratefully acknowledge the scholarships I received, the ANU scholarship, the Olin-Eggen scholarship, the RSAA supplementary scholarship, and the funding by the Australian Research Council Discovery Project, DP140103341 \textit{The key role of black holes in galaxy formation}. 

The computations were undertaken on the National Computational Infrastructure located at the Australian National University. 

Finally, to my parents who raised me with selfless love and so many sacrifices.





